---
title: "RepData Assignment_1.Rmd"
output: html_document
description: R Markdown document for Coursera course "Reproducible Research", Peer Assessment 1
---
# Reproducible Research: Peer Assessment 1
R Markdown document for Reproducible Research Peer Assessment 1

## Loading and preprocessing the data
```{r preprocessing}
setwd("~/DS Toolbox/Github/RepData_PeerAssessment1")
activity <- read.csv("activity.csv")  # read file
#activity$date <- as.Date(activity$date, "%Y-%m-%d")  # convert date column to date data type
```

## What is mean total number of steps taken per day?
1. Histogram with number of steps for each day
```{r dailysteps_hist}
dailysteps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)  # calculate daily steps
# library(lattice)
# histogram(dailysteps, type="count", breaks=length(dailysteps))  # output similar to using ggplot2
library(ggplot2)
qplot(dailysteps)
```

2. **mean** and **median** total number of steps taken per day
```{r mean_median}
mean(dailysteps)  # mean
median(dailysteps)  # median
summary(dailysteps)  # different approach
```

## What is the average daily activity pattern?
1. Time-series plot of the 5-minute interval and average number of steps taken
```{r interval_average}
intervalsteps <- tapply(activity$steps, factor(activity$interval)
                            , mean, na.rm=TRUE)  # 5 min interval avg
# if using lattice library
# xyplot(intervalsteps ~ factor(activity$interval), type="l"
#            , xlab = "5-min interval", ylab = "Number of Steps"
#            , scales = list(x = list(tick.number = 24)))

# easy plot using ggplot2
qplot(as.numeric(names(intervalsteps)), intervalsteps
            , geom="line", xlab="Time of Day HHMM", ylab="Average Steps")
```

2. Identify the 5-min interval that contains the maximum number of steps
```{r max_daily_steps}
moststeps <- order(intervalsteps, decreasing=TRUE)  # daily average in decreasing order
intervalsteps[moststeps[1]]  # interval with maximum number of steps
```

## Inputting missing values
1. Total number of missing values in the dataset
```{r total_na}
sum(is.na(activity$steps))
```
2. Fill-in missing values in the dataset
```{r na_filler}

```

## Are there differences in activity patterns between weekdays and weekends?

