---
title: "RepData Assignment_1.Rmd"
output: html_document
description: R Markdown document for Coursera course "Reproducible Research", Peer Assessment 1
---
# Reproducible Research: Peer Assessment 1
R Markdown document for Reproducible Research Peer Assessment 1

## Loading and preprocessing the data
```{r preprocessing}
setwd("~/DS Toolbox/Github/RepData_PeerAssessment1")
activity <- read.csv("activity.csv")  # read file
#activity$date <- as.Date(activity$date, "%Y-%m-%d")  # convert date column to date data type
```

## What is mean total number of steps taken per day?
1. Histogram with number of steps for each day
```{r dailysteps_hist}
dailysteps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)  # calculate daily steps
library(lattice)
histogram(dailysteps, type="count", breaks=length(dailysteps))
```

2. **mean** and **median** total number of steps taken per day
```{r mean_median}
mean(dailysteps)  # mean
median(dailysteps)  # median
summary(dailysteps)  # different approach
```

## What is the average daily activity pattern?
1. Time-series plot of the 5-minute interval and average number of steps taken
```{r interval_average}
intervalsteps <- tapply(activity$steps, factor(activity$interval)
                            , mean, na.rm=TRUE)  # 5 min interval avg
# code for timeseries plot
```

2. Identify the 5-min interval that contains the maximum number of steps
```{r max_daily_steps}
moststeps <- order(intervalsteps, decreasing=TRUE)  # daily average in decreasing order
intervalsteps[moststeps[1]]  # interval with maximum number of steps
```

## Inputing missing values

## Are there differences in activity patterns between weekdays and weekends?

